### 2023-09-07 draft.tex
adding stuff

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  11pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}
\usepackage{float}
\usepackage{mathtools}
\usepackage{natbib}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\setcitestyle{numbers,square,comma}
\usepackage{verbatim}
\usepackage{amsthm}
\usepackage{comment}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{\vspace{-2.5em}}

\begin{document}

\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
\newcommand{\SBM}{\mathrm{SBM}}
\newcommand{\DCBM}{\mathrm{DCBM}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\MBM}{\mathrm{MBM}}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{simulation-study}{%
\section{Simulation Study}\label{simulation-study}}

\hypertarget{classification}{%
\subsection{Classification}\label{classification}}

In this simulation experiment, the latent vectors were sampled along a
Bezier curve defined by
\(g(t) = \begin{bmatrix} t^2 & 2 t (1-t) \end{bmatrix}^\top\). The
timepoints \(t_i\) were sampled as iid Beta random variables with two
sets of parameters, \((\alpha_1, \beta_1)\) and \((\alpha_2, \beta_2)\).
The setup is as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Draw response variables \(y_1, ..., y_n \iid \Multinomial(1/2, 1/2)\).
\item
  For each \(i = 1, ..., n\), draw
  \(t_i \mid y_i \indep \Betadist(\alpha_{y_i}, \beta_{y_i})\), such
  that

  \begin{itemize}
  \tightlist
  \item
    \(\alpha_1 = 1\)
  \item
    \(\beta_1 = 2\)
  \item
    \(\alpha_2 = 2\)
  \item
    \(\beta_2 = 1\)
  \end{itemize}
\item
  Construct each latent vector as \(x_i = g(t_i)\) and compile them in
  data matrix
  \(X = \begin{bmatrix} x_1 & \cdots & x_n \end{bmatrix}^\top\).
\item
  Sample graph and its adjacency matrix as \(A \sim \RDPG(X)\).
\end{enumerate}

For each graph, we constructed the ASE, which was used to estimate the
parameters \((\hat{\alpha}, \hat{\beta})\) for the graph, using the
maximum likelihood method. The estimated parameters were then used as
predictors \(y_1, ..., y_n\), setting aside half for training and half
for testing. We investigated graphs of size \(|V| = 32, 64, 128, 256\).
The number of graphs for each experiment was set to \(n = 64\). For each
(number of graphs, size of graph) pair, we performed 32 replicates.
Figure \ref{fig:class-sim-ase} shows the ASE of one graph.\\
Figure \ref{fig:classification-sim} shows the boxplots of the
classification error rates.

\begin{center}\includegraphics{draft_files/figure-latex/class-sim-ase-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/classification-sim-1} \end{center}

\hypertarget{regression}{%
\subsection{Regression}\label{regression}}

\hypertarget{applications}{%
\section{Applications}\label{applications}}

HCP data

\begin{center}\includegraphics{draft_files/figure-latex/hcp-eda-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/hcp-gender-1} \end{center}

\begin{center}\includegraphics{draft_files/figure-latex/hcp-age-1} \end{center}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\renewcommand\refname{Bibliography}
  \bibliography{bibliography.bib}

\end{document}